% $State: Exp $ $Date: 2003/01/30 11:25:53 $ $Revision: 1.3 $

\chapter{Flags index.}\label{flag}

In this index we list all the \AFD\ flags (see the description of the command
{\tt flag} in the index \ref{cmd} to learn how to print and modify the value of
these flags).

\begin{description}
  \item [\tt auto\_lvl] (integer, default is 0) : Control the
  automatic detection of axioms: If it is set to 0 no detection is
  performed. If it is set to 1, axioms are detected when the goal is
  structurally equal to an hypothesis. If it is set to 2, axioms are
  detected when the goal unifies is equal to an hypothesis up to the
  expansion of some definitions. If it is set to 3, axioms are
  detected if the goal unifies with an hypothesis (using no
  equations). We recommend avoiding 3 as it may instantiate variables
  with the wrong value.

  \item [\tt auto\_type] (bool, default is false) : automatically
apply all the introduction rule which were introduced with the flags
{\tt -i} and {\tt -c} or {\tt -t}. We recommend setting this flag to
true and using {\tt auto\_lvl} set to 2 to solve automatically all the
``typing'' goals (like proving that something is an integer).

  \item [\tt binder\_tex\_space] (integer, default is 3) : set the space after
    a binder when \AFD is printing TeX formulas.

  \item [\tt comma\_tex\_space] (integer, default is 5) : set the space after
    punctuation when \AFD is printing TeX formulas.

  \item [\tt ellipsis\_text] (string, default is ``...'') : the text to be
    printed when an expression is too deep (used by the pretty printer only).

  \item [\tt eq\_breadth] (integer, default is 4) : maximum number of
    equations used at each step of rewriting.

  \item [\tt eq\_flvl] (integer, default is 3) : maximum number of
    interleaved equations tried without decreasing the distance (the
    rewriting algorithm uses a distance between first order terms).

  \item [\tt eq\_depth] (integer, default is 100) : maximum number of
    interleaved equations applied by the rewriting algorithm.

  \item [\tt margin] (integer, default is 80) : size of the page (used by the
    pretty printer only).

  \item [\tt max\_indent] (integer, default is 50) : maximum number of
    indentation (used by the pretty printer only).

  \item [\tt max\_boxes] (integer, default is 100) : control the maximum
    printing ``depth''. If the expression is too deep, an ellipsis is printed
    (used by the pretty printer only).

  \item [\tt min\_tex\_space] (integer, default is 20) : set the minimum space
    (in 100th of em) between to tokens when \AFD\ is printing TeX formulas.

  \item [\tt max\_tex\_space] (integer, default is 40) : set the minimum space
    (in 100th of em) between to tokens when \AFD\ is printing TeX formulas.

  \item [\tt tex\_indent] (integer, default is 200) : set the indentation space
    (in 100th of em) used by \AFD\ when printing multi-lines TeX formulas.

  \item [\tt tex\_lisp\_app] (boolean, default is true) : If true the syntax
     $f\;x\;y$ is used for application when producing \LaTeX\ formulas. If
     false, the syntax $f(x,y)$ is used.

  \item [\tt tex\_type\_sugar] (boolean, default is true) : If true the
    syntactic sugar $\forall x:A\;B$ for $\forall x (A x \to B)$ is
    used  when producing \LaTeX\ formulas.

  \item [\tt tex\_margin] (integer, default is 80) : size of the page used
    when printing verbatim formulas in TeX.

  \item [\tt tex\_max\_indent] (integer, default is 50) : maximum number of
    indentation used when printing verbatim formulas in TeX.

  \item [\tt trivial\_depth] (integer, default is 4) : default value for the
    {\tt trivial} command.

\end{description}


